Carbonate chemistry at Ny-Ålesund

### 2022-08-16 awipev-CO2.Rmd
Added fCO2 calculated at REMOS depths

# we look at pCO2 normalized at 4°C at all depths
# calculate fCO2 from pH and salinity-derived AT
  dplyr::select(datetime, depth, pHint_tot_sf, sal_insitu_ctd, temp_insitu_ctd) %>% 
  drop_na() %>% 
  dplyr::mutate(at_0_4m = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
                pco2_0_4m = seacarb::carb(flag=8, var1=pHint_tot_sf, var2=tmp$at_0_4m)$pCO2)
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pHint_tot_sf_temp_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf at 4°C vs depth", x = "Depth (m)", y= "pH at 4°C (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group * tmp$montht))

pH_delta_plt <- tidyr::drop_na(tmp_delta) %>% 
  ggplot(aes(x=montht, y=pH_tot_sf_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(substitute("pH"["T"] ~ "at 4°C"), limits = c(-0.2, 0.2), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 1.5, y = 0.8, parse = TRUE, label = "pH['surface'] > pH['deep']") +
  annotate("text", x = 1.5, y = -0.15, parse = TRUE, label = "pH['surface'] < pH['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
# all deltas
deltas_plt <- sal_delta_plt / temp_delta_plt / den_delta_plt / pH_delta_plt
deltas_plt <- deltas_plt + plot_annotation(tag_levels = 'a')
print(deltas_plt)
ggsave(filename = "figures/deltas.png", plot = deltas_plt, width = 15, height = 20, units = "cm")
```

@@

seaFET pH is significantly different at the surface and depth. It should
therefore be the same for pCO2.

Let's look at the stratification tab above (salinity, temperature, density and pH).

## pCO2 from pH and AT 0-4m
If one calculates pCO2 from seafet pH and AT (from salinity) between 0-4 m, do we get better estimates of air-sea CO2 fluxes?

```{r}
size_point <- 0.3 # size of data points
alpha <- 1 #transparency
size_labs <- 6 # font size labels

tmp <- d_all %>% 
  dplyr::select(datetime, pHint_tot_sf, sal_insitu_ctd, temp_insitu_ctd) %>% 
  drop_na() %>% 
  dplyr::mutate(at_0_4m = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
                pco2_0_4m = seacarb::carb(flag=8, var1=pHint_tot_sf, var2=tmp$at_0_4m)$pCO2)
ts_pCO2_calc <- tmp %>%
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pco2_0_4m), 
             colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title=expression(paste(pCO[2], "calc at 0-4 m and in situ temperature")),
       x="", y=TeX("pCO$_2$ ($\\mu atm$)")) +
  Mytheme(size_labs = size_labs)
ts_pCO2_calc
```

>>>>

# we look at pCO2  at all depths
# calculate fCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_sf = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    fco2_sf = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_sf,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$fCO2
  ) %>% 
  dplyr::select(datetime, fco2_sf)
d_all <- left_join(d_all, tmp, by = "datetime")
tmp <- d_all %>% 
  dplyr::select(montht, fco2_sf, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = fco2_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "fco2_sf vs depth", x = "Depth (m)", y= "fCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$fco2_sf ~ tmp$depth_group * tmp$montht))

# we look at fCO2 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, fco2_sf, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$fco2_sf ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(fco2_sf_med = median(fco2_sf, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = fco2_sf_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%   dplyr::mutate(fco2_sf_med_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 3, caption = "Median seafet fCO2 with (median0_4m - median8_12m)")

p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group2), y = fco2_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pH at 4°C vs depth", x = "Depth (m)", y= "pHint_tot_sf") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

fco2_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = fco2_sf_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(substitute("fCO"["2"]), limits = c(-45, 10)) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 2.5,
    y = 5,
    parse = TRUE,
    label = "fCO['2 surface'] > fCO['2 deep']"
  ) +
  annotate(
    "text",
    x = 2.5,
    y = -40,
    parse = TRUE,
    label = "fCO['2 surface'] < fCO['2 deep']"
  ) +
  Mytheme(size_labs = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
fco2_delta_plt
# all deltas
deltas_plt <- sal_delta_plt / temp_delta_plt / den_delta_plt / fco2_delta_plt
deltas_plt <- deltas_plt + plot_annotation(tag_levels = 'a')
print(deltas_plt)
ggsave(filename = "figures/deltas.png", plot = deltas_plt, width = 15, height = 20, units = "cm")
```

@@

Let's look at the stratification tab above (salinity, temperature, density and fCO2): fCO2 is generally higher in the bottom layer than in the surface layer (note that no data is available in May, June and July). Which means that the air-sea CO2 fluxes calculated with pCO2 at 11 m are overestimated.
